Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-07-06 paper_jasa.tex
cleaned section 6.7

In the previous section we noticed that subjects tend to pick the plot with the lowest $p$-value. If a null plot has data that, by chance because the sample is a random draw from the null, has a smaller $p$-value than the actual data, we would expect that there is a strong possibility that an subject will pick this null plot instead of the actual data plot. Figure \ref{fig:pval_difference} investigates this for the data collected in experiments 1 and 2. The plots display the proportion of subjects who choose the actual data plot against the $p$-value of the actual data minus the lowest seen in the null data. A small negative value of the difference means that the $p$-value of the actual data is smaller than that of the null data. If there is a big negative difference we'd expect to see that the subjects would more often make the correct choice of plot. When there is a large positive difference, subjects fail to pick the correct plot. If however, there is little difference between the $p$-values then we'd expect to see more confusion in the choices.

\green{Something is definitely not right! The code looks like it does actual data p-value - min(null $p$-value). But how can you get a difference of 0.8 from this??} \blue{It is possible theoretically! Suppose data plot $p$-value is larger than 0.8. Under null hypothesis, $p$-value $\sim U(0,1)$. Thus min(null $p$-value) could be close to zero by chance producing this large difference. And this has happened in both experiment 1 and 2.}

 \green{Is this really pct correct, or 1-pct correct? \blue{This is proportion correct}. When there is a big diff in p-value almost no-one gets it right! Maybe it is just a scale issue?} \blue{In fact this is what we expect. If there is large difference that means data plot $p$-value is large while minimum $p$-value is very small. This should lead people to pick the wrong plot. Our minimum $p$-value assumption basically says this fact, right?}. We see that when there is a big difference between the $p$-values subjects choose the actual data plot more often.

>>>>

In the previous section we noticed that subjects tend to pick the plot with the lowest $p$-value. If a null plot has data that, by chance because the sample is a random draw from the null, has a smaller $p$-value than the actual data, we would expect that there is a strong possibility that an subject will pick this null plot instead of the actual data plot. Figure \ref{fig:pval_difference} investigates this for the data collected in experiments 1 and 2. The plots display the proportion of subjects who choose the actual data plot against the $p$-value of the actual data minus the lowest seen in the null data. A small negative value of the difference means that the $p$-value of the actual data is smaller than that of the null data. If there is a big negative difference we'd expect to see that the subjects would more often make the correct choice of plot. When there is a large positive difference, subjects fail to pick the correct plot. If however, there is little difference between the $p$-values then we'd expect to see more confusion in the choices. That is why we see a big range of proportion correct where difference is close to zero in Figure \ref{fig:pval_difference}.

%\green{Something is definitely not right! The code looks like it does actual data p-value - min(null $p$-value). But how can you get a difference of 0.8 from this??} \blue{It is possible theoretically! Suppose data plot $p$-value is larger than 0.8. Under null hypothesis, $p$-value $\sim U(0,1)$. Thus min(null $p$-value) could be close to zero by chance producing this large difference. And this has happened in both experiment 1 and 2.}

 %\green{Is this really pct correct, or 1-pct correct? \blue{This is proportion correct}. When there is a big diff in p-value almost no-one gets it right! Maybe it is just a scale issue?} \blue{In fact this is what we expect. If there is large difference that means data plot $p$-value is large while minimum $p$-value is very small. This should lead people to pick the wrong plot. Our minimum $p$-value assumption basically says this fact, right?}. We see that when there is a big difference between the $p$-values subjects choose the actual data plot more often.
